\documentclass[12pt]{article}
\pagestyle{empty}

\usepackage{graphicx}
\usepackage{float}
\usepackage{times}
\usepackage{color}
\usepackage{listings}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{ascii}
\usepackage[T1]{fontenc}
\usepackage{datetime}
\usepackage[top=1.5in, bottom=1.5in, left=1in, right=1in]{geometry}

\usepackage{algorithm2e}
  
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\pagestyle{fancy} 
\renewcommand{\headrulewidth}{0pt}
\shortdate
\fancyhead[LE, RO]{\textit{\date{\currenttime} \today}}
\fancyhead[LO]{\textit {Xiaofan Lu}}



\titleformat{\section}{\bf\Large\centering}{\thesection}{1em}{}
  
\begin{document}
\title{CS388L Quiz 4}
\author{Xiaofan Lu, xl4326}
\date{\vspace{-3ex}}
\maketitle

\noindent
Let $\Gamma$ be a positive program containing the rule $p \leftarrow q$. Show that if $p$ doesn't occur in the heads of the other rules of $\Gamma$ then every minimal model of $\Gamma$ satisfies the formula $p \leftrightarrow q$. 

\begin{proof}
For any minimal model $M$ of $\Gamma$, 
\begin{itemize}
   \item If $q \in M$, since the rule $p \leftarrow q$ must be satisfied by $M$, then $p \in M$ as well. Clearly, M satisfies $p \leftrightarrow q$.  
   \item If $q \not \in M$, the rule $p \leftarrow q$ is satisfied. Suppose $p \in M$, define $M' = M \backslash \{p\}$. Clearly, the rule $p \leftarrow q$ is satisfied by $M'$ as well. For all other rules $H \leftarrow B$ in $\Gamma$:
   \begin{itemize}
       \item If $M$ satisfies $H$, since $p$ doesn't occur in $H$, $M'$ satisfies $H$ as well. Then $M'$ satisfies this rule. 
       \item If $M$ doesn't satisfy $H$, since $M$ satisfies $H \leftarrow B$, $M$ must not satisfy $B$. By problem 24, $M'$, a subset of $M$, doesn't satisfy $B$ as well. Thus, $M'$ also satisfies this rule. 
   \end{itemize}
   This is to say, $M'$ satisfies $\Gamma$, which contradicts that $M$ is a minimal model of $\Gamma$. Thus, our assumption that $p \in M$ must be false. This is to say, if $q \not \in M$, $p \not \in M$ as well. Clearly, M satisfies $p \leftrightarrow q$. 
\end{itemize}
To sum up, if $p$ doesn't occur in the heads of the other rules of $\Gamma$ then every minimal model of $\Gamma$ satisfies the formula $p \leftrightarrow q$. 

\end{proof}

\end{document}























